---
title: "README"
author: "Sebastian Hellmann and Manuel Rausch"
date: "2024-04-10"
output:
  md_document:
    variant: gfm
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
gitrep <- "https://github.com/ManuelRausch/StatConfR/tree"
gitbranch <- "main/"
```

# statConfR: Models of Decision Confidence and Metacognition

The statConfR package provides functions to fit static models of
decision-making and confidence derived from signal detection theory for
binary discrimination tasks, as well as meta-d′/d′ (Rausch & Hellmann,
2024). The package can be used to test the assumptions underlying
meta-d′/d′. Several models provide a metacognition parameter that may
serve as an alternative when the assumptions of meta-d′/d′ assuming the
corresponding model provides a better fit to the data. Currently, the
following models are included:

-   Signal detection rating model

-   Gaussian noise model

-   Weighted evidence and visibility model

-   Post-decisional Accumulation Model

-   Independent Gaussian model

-   Independent truncated gaussian Model (the model underlying the
    meta-d′/d′ method, see Rausch et al., 2023)

-   Lognormal noise model

-   Lognormal weighted evidence and visibility model

## Installation

For the current development version, the easiest way of installation is
using `devtools` and install from GitHub:

```         
devtools::install_github("ManuelRausch/StatConfR")
```

The latest released version of the package will soon be available on
CRAN via

``` r
install.packages("statConfR")
```

## Usage

### Data structure

The package includes a demo data set from a masked orientation
discrimination task with confidence judgments (Hellmann et al., 2023,
Exp. 1.

```{r}
library(statConfR)
data("MaskOri")
head(MaskOri)
```

Data should be in the form of a data.frame object columns for following
variables:

-   stimulus (factor with 2 levels): The property of the stimulus which
    defines which response is correct
-   diffCond (factor): The experimental manipulation that is expected to
    affect discrimination sensitivity
-   correct (0-1): Indicating whether the choice was correct (1) or
    incorrect(0).
-   rating (factor): A discrete variable encoding the decision
    confidence (high: very confident; low: less confident)
-   participant (integer): giving the subject ID.

### Fitting

It is strongly recommended that if metacognitive efficiency is to be
measured using the meta-d′/d′ method that researchers fist determine
whether the Independent Truncated Gaussian Model, the confidence model
implied by the meta-d′/d′ method, is an adequate description of the
data. Using the function fitConfModel, we can fit several confidence
models to the data of each participant. The argument
`.parallel=TRUE`allows for parallelization over all but one available
core.

```{r}
fitted_pars <- fitConfModels(MaskOri, models=c("SDT", "WEV"), .parallel = TRUE) 
```

This parallelizes the fitting process over participant-model
combinations. The output is then a data frame with one row for each
participant-model combination and columns for parameters and measures
for model performance (negative log-likelihood, BIC, AIC and AICc).
These may be used for quantitative model comparison.

```{r, echo=FALSE}
fitted_pars <- structure(list(
  model = c("SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV",             "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV",             "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV",             "SDT", "WEV"), 
  participant = c(1, 1, 2, 2, 3, 3, 4, 4, 
                  5, 5, 6, 6, 7, 7, 8, 8, 9, 9, 10, 10, 11, 11, 12, 12, 13, 13, 
                  14, 14, 15, 15, 16, 16), 
  negLogLik = c(2721.256, 2621.110, 1946.258, 1827.221, 1706.178, 
                1661.617, 1315.112, 1247.284, 1820.819, 1723.671,
                2593.780, 2514.932, 1698.055, 1612.056, 2544.435, 
                2491.750, 2347.679, 2219.598, 1715.222, 1582.766, 2182.176,                      1868.587, 2485.889, 2393.366, 1096.712, 1048.117, 2030.395,                      1821.567, 2306.528, 2156.135, 2434.188, 2193.127), 
  N = c(1620L, 1620L, 1620L, 1620L, 1620L, 1620L, 1620L, 1620L, 
        1620L, 1620L, 1620L, 1620L, 1620L, 1620L, 1620L, 1620L, 
        1620L, 1620L, 1620L, 1620L, 1620L, 1620L, 1620L, 1620L, 
        1620L, 1620L, 1620L, 1620L, 1620L, 1620L, 1620L, 1620L), 
  k = c(14L, 16L, 14L, 16L, 14L, 16L, 14L, 16L, 
        14L, 16L, 14L, 16L, 14L, 16L, 14L, 16L, 
        14L, 16L, 14L, 16L, 14L, 16L, 14L, 16L, 
        14L, 16L, 14L, 16L, 14L, 16L, 14L, 16L), 
  BIC = c(5545.975, 5360.464, 3995.979, 3772.684, 
          3515.818, 3441.476, 2733.686, 2612.811, 
          3745.101, 3565.585, 5291.023, 5148.107, 
          3499.572, 3342.356, 5192.333, 5101.744, 
          4798.820, 4557.439, 3533.907, 3283.774,
          4467.814, 3855.416, 5075.241, 4904.975, 
          2296.886, 2214.477, 4164.253, 3761.376,
          4716.518, 4430.512, 4971.838, 4504.496), 
  AICc = c( 5470.739, 5274.520, 3920.743, 3686.741,
            3440.582, 3355.533, 2658.451, 2526.868,
            3669.866, 3479.642, 5215.787, 5062.164,
            3424.336, 3256.412, 5117.097, 5015.800, 
            4723.584, 4471.495, 3458.671, 3197.831, 
            4392.578, 3769.473, 5000.005, 4819.031,
            2221.650, 2128.533, 4089.017, 3675.432, 
            4641.282, 4344.569, 4896.602, 4418.553), 
  AIC = c(5470.513, 5274.221, 3920.517, 3686.441, 
          3440.356, 3355.233, 2658.224, 2526.569, 
          3669.639, 3479.342, 5215.560, 5061.865,
          3424.110, 3256.113, 5116.870, 5015.501,
          4723.358, 4471.196, 3458.444, 3197.531,
          4392.351, 3769.174, 4999.779, 4818.732, 
          2221.423, 2128.234, 4088.790, 3675.133,
          4641.055, 4344.269, 4896.375, 4418.253), 
  d_1 = c(0.0428, 0.2027, 0.0000, 0.0512, 0.2708, 0.4146, 0.0002, 0.0266,
          0.0001, 0.0002, 0.0070, 0.0094, 0.0133, 0.1092, 0.1272, 0.0283, 
          0.0874, 0.1894, 0.0979, 0.0815, 0.0036, 0.0000, 0.0394, 0.2394, 
          0.0014, 0.0008, 0.0371, 0.0029, 0.0506, 0.2515, 0.0043, 0.0056), 
  d_2 = c(0.4593, 0.6142, 0.0950, 0.1920, 0.4673, 0.8561, 0.0785, 0.1828, 
          0.1510, 0.1708, 0.3743, 0.3872, 0.5764, 0.6861, 0.1283, 0.1393, 
          0.2870, 0.4464, 0.2501, 0.2981, 0.0048, 0.0000, 0.3578, 0.2691,
          0.0288, 0.0030, 0.0553, 0.0494, 0.3503, 0.5728, 0.2345, 0.2263), 
  d_3 = c(1.0526, 1.0797, 0.8601, 1.0412, 1.9117, 2.7115, 2.6508, 2.2663, 
          1.5160, 1.5043, 0.8712, 1.2054, 1.7979, 1.8298, 0.5464, 0.7973, 
          1.0209, 1.0596, 0.5550, 0.6764, 0.0067, 0.2912, 0.6599, 0.8339, 
          0.5883, 0.7724, 2.6441, 2.1897, 1.8880, 1.9333, 0.3032, 0.5420), 
  d_4 = c(3.6806, 3.4746, 6.1410, 4.1423, 6.4257, 6.9164, 7.3334, 4.7879, 
          5.8306, 3.7159, 4.0074, 3.2443, 5.9603, 4.5527, 2.6227, 2.5127, 
          5.2454, 3.7840, 3.1460, 2.3962, 2.4328, 1.9491, 2.7917, 2.5463, 
          6.3653, 4.5931, 6.8276, 4.2693, 5.3671, 4.0775, 1.6437, 1.8035), 
  d_5 = c(4.7779, 4.0799, 8.0556, 5.2886, 7.5755, 7.9863, 8.3112, 5.3117, 
          7.5051, 4.5348, 6.3132, 4.6611, 8.6092, 6.2114, 5.3301, 4.6658, 
          9.1885, 5.8474, 6.0059, 3.3914, 6.6479, 3.3889, 6.0167, 4.4172, 
          9.2753, 6.4803, 9.1068, 5.3490, 7.7549, 5.2797, 5.4378, 2.8724), 
  c = c(-0.2723, -0.2957, -0.1394, -0.1475, -1.1510, -1.3743, 0.1900, 0.1717,            -0.1358, -0.1179, -0.1861, -0.1905, -0.1001, -0.1248, 0.4126, 0.4138, 
        -0.2692, -0.2673, -0.0105, -0.0008, -0.1961, -0.1895, -0.1925, -0.2076,          -0.9455, -0.9123, -0.2795, -0.2606, -0.1878, -0.1973, 0.1028,  0.1013),
  theta_minus.4 = c(-1.5467, -2.0665, -2.0092, -2.0441,
                    -1.9938, -2.7625, -2.4023, -2.3219, 
                    -2.4831, -2.2074, -2.0534, -2.3283, 
                    -1.9612, -2.2717, -1.7405, -2.2854, 
                    -2.5484, -2.6862, -5.1218, -3.7997, 
                    -3.7311, -2.8340, -1.4923, -1.7448, 
                    -2.6828, -2.5758, -3.4180, -3.0854, 
                    -1.8803, -2.1973, -1.4694, -1.5659), 
  theta_minus.3 = c(-1.0333, -1.2485, -1.9193, -1.9500, 
                    -1.6372, -1.9192, -2.1014, -2.0707, 
                    -1.7127, -1.6231, 1.4491, -1.6752, 
                    -1.7638, -2.0425, -1.5328, -2.0199, 
                    -1.7582, -1.8359, -4.4336, -3.3743, 
                    -2.2426, -1.8941, -1.2401, -1.4400, 
                    -2.3699, -2.3140, -2.0691, -2.0793, 
                    -1.2779, -1.5282, -1.3487, -1.4590), 
  theta_minus.2 = c(-0.6336, -0.4152, -1.4097, -1.3982, -1.2600, -0.3724,                            -1.9896, -1.9636, -1.5134, -1.4513, -1.1293, -1.2743,                            -1.7095, -1.9781, -1.2594, -1.6607, -1.1097, -1.1141,                            -3.3155, -2.6892, -1.8550, -1.5981, -1.0955, -1.2453,                            2.2514, -2.2057, -1.6390, -1.6791, -0.9313, -1.0813,                            -1.0651, -1.1707), 
  theta_minus.1 = c(-0.4543,  0.1296, -0.9580, -0.9030, -1.1668, 0.9328,
                    -1.7341, -1.7092, -1.2946, -1.2548, -0.6609, -0.5921,
                    -1.2117, -1.3993, -0.7776, -1.0002, -0.5582, -0.3386,
                    -1.7153, -1.6536, -1.4748, -1.2134, -0.7674, -0.7641,
                    -2.0412, -1.9852, -0.9709, -0.9152, -0.4497, -0.2818,
                    -0.9105, -0.9926), 
  theta_plus.1 = c(-0.0944, -0.6196, 0.7857,  0.8201, -1.1143, -2.7695,  
                   2.0205,  2.0109,  1.5262,  1.3854,  0.5231,  0.5995,
                   1.2145, 1.5088, 0.9446,  0.8995 , 0.1476,  0.1797,
                   1.7219,  1.6238,  1.0330,  0.9896, 0.7259,  0.8791,  
                   1.8260,  1.6742,  0.9272,  0.9170,  0.1505,  0.1830,  
                   0.9580,  0.8831), 
  theta_plus.2 = c(0.2152,  0.1544,  1.3781,  1.4484, -0.7344, -1.1313,  
                   2.2051,  2.1714,  1.7948,  1.6133,  1.0471,  1.2389,  
                   1.6436,  1.9894,  1.2760,  1.5454,  0.8249,  0.9612,  
                   2.9214,  2.4354,  1.6332,  1.5325,  1.3923,  1.6567, 
                   2.1620,  2.0029,  1.5213,  1.5131,  0.8220,  1.0777,  
                   1.1148,  1.0462), 
  theta_plus.3 = c(0.9850, 1.3976, 2.0879, 2.2447, 0.2961, 0.7714, 
                   2.3807, 2.3164, 2.0180, 1.7985, 1.3480, 1.5777, 
                   1.8176, 2.1864, 1.6817, 2.1866, 1.3751, 1.5394, 
                   4.1369, 3.1720, 2.0967, 1.8586, 1.5691, 1.8518, 
                   2.3198, 2.1595, 2.1597, 2.0872, 1.1492, 1.4735, 
                   1.7962, 1.7024), 
  theta_plus.4 = c(1.5735, 2.1879, 2.2369, 2.4030, 0.9314, 1.7520, 
                   2.7241, 2.5805, 2.9750, 2.5276, 1.9832, 2.2356,
                   2.1917, 2.5857, 1.9370, 2.5482, 2.3346, 2.5794, 
                   5.2679, 3.8478, 3.8224, 2.8743, 1.8500, 2.1539, 
                   2.7041, 2.5091, 3.7390, 3.3099, 1.8606, 2.2707, 
                   1.8980, 1.7892),
  sigma = c(NA, 1.0105, NA, 0.6391, NA, 1.3289, NA, 0.3921, 
            NA, 0.2973, NA, 0.7526, NA, 0.7537, NA, 1.0749, 
            NA, 0.6463, NA, 0.2883, NA, 0.3090, NA, 0.7764, 
            NA, 0.3878, NA, 0.4180, NA, 0.6858, NA, 0.5179), 
  w = c(NA, 0.5361, NA, 0.5020, NA, 0.3818, NA, 0.3805,
        NA, 0.3596, NA, 0.4417, NA, 0.4783, NA, 0.4843,
        NA, 0.5162, NA, 0.4645, NA, 0.5261, NA, 0.4832,
        NA, 0.2917, NA, 0.4503, NA, 0.5133, NA, 0.6329), 
  wAIC = c(0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 
           0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1), 
  wAICc = c(0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 
            0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1), 
  wBIC = c(0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 
           0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1, 0, 1)), class = "data.frame", row.names = c(NA, 
                                                                                           -32L))
```

```{r}
head(fitted_pars)
```

If the Independent Truncated Gaussian model provides a decent account of the data (which is not the case though in the demo dataset), it is legitimate to quantify metacognitive efficiency with meta-d′/d′:

``` {r}
MetaDs <- fitMetaDprime(subset(MaskOri, diffCond == "33.3"), 
                        model="ML", .parallel = TRUE)
```

## Contact

For comments, remarks, and questions please contact either
[manuel.rausch\@hochschule-rhein-waal.de](mailto:manuel.rausch@hochschule-rhein-waal.de)
or [sebastian.hellmann\@ku.de](mailto:sebastian.hellmann@ku.de){.email}
or [submit an issue](https://github.com/ManuelRausch/StatConfR/issues).

## References

Hellmann, S., Zehetleitner, M., & Rausch, M. (2023). Simultaneous
modeling of choice, confidence, and response time in visual perception.
Psychological Review. 130(6), 1521--1543.
[doi:10.1037/rev0000411](https://doi.org/10.1037/rev0000411)

Rausch, M., Hellmann, S. & Zehetleitner, M. (2023). Measures of
metacognitive efficiency across cognitive models of decision confidence.
Psychological Methods.
[doi:10.1037/met0000634](https://doi.org/10.1037/met0000634)

Rausch, M., & Hellmann, S. (2024). statConfR: An R Package for Static
Models of Decision Confidence and Metacognition. PsyArXiv.
[doi:10.31234/osf.io/dk6mr](https://doi.org/10.31234/osf.io/dk6mr)
