---
title: "README"
author: "Manuel Rausch, Sascha Meyen, and Sebastian Hellmann"
date: "2024-10-13"
output:
  md_document:
    variant: gfm
editor_options: 
  markdown: 
    wrap: 72
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
gitrep <- "https://github.com/ManuelRausch/StatConfR/tree"
gitbranch <- "main/"
```


The `statConfR` package provides functions to fit static models of
decision-making and confidence derived from signal detection theory for
binary discrimination tasks. Up to now, the following models have been included:

- signal detection rating model (Green & Swets, 1966),
- Gaussian noise model (Maniscalco & Lau, 2016),
- weighted evidence and visibility model (Rausch et al., 2018),
- post-decisional accumulation model (Rausch et al., 2018),
- independent Gaussian model (Rausch & Zehetleitner, 2017),
- independent truncated Gaussian model (the model underlying the meta-d′/d′ method, see Rausch et al., 2023),
- lognormal noise model (Shekhar & Rahnev, 2021), and
- lognormal weighted evidence and visibility model (Shekhar & Rahnev, 2023).

Bayesian model selection on the group level is performed using a fixed effects model (i.e. assuming that data from each subject was caused by the same generative model) and a and a Dirichelet random-effects model (which assumes that different generative models may cause the data from different subjects), as proposed by Rigoux et al. (2014) and Daunizeau et al. (2014). 

In addition, the `statConfR` package provides functions for estimating different 
kinds of measures of metacognition: 
- meta-d$^\prime$/d$^\prime$, the most widely-used measure of metacognitive efficiency, allowing both Maniscalco and Lau (2012)'s and Fleming (2017)'s model specification. Fitting models of confidence is a way to test the assumptions underlying meta-d′/d′. 
- Information-theoretic measures of metacognition (Dayan, 2023), including
  - meta-I, an information-theoretic measures of metacognitive sensitivity, 
  - $meta-I_{1}^{r}$ and $meta-I_{2}^{r}$, two measures of metacognitive efficiency proposed by Dayan (2023), 
  - RMI, a novel measure of metacognitive accuracy, also derived from information theory.  
  


## Mathematical description of implemented models of confidence
The models included in the statConfR package are all based on signal detection theory (Green & Swets, 1966). 
It is assumed that participants select a binary discrimination response $R$ about a stimulus $S$.
Both $S$ and $R$ can be either -1 or 1. $R$ is considered correct if $S=R$.
In addition, we assume that in the experiment, there are $K$ different levels of stimulus discriminability, 
i.e. a physical variable that makes the discrimination task easier or harder.
For each level of discriminability, the function fits a different discrimination
sensitivity parameter $d_k$. If there is more than one sensitivity parameter,
we assume that the sensitivity parameters are ordered such as $0 < d_1 < d_2 < ... < d_K$.
The models assume that the stimulus generates normally distributed sensory evidence $x$ with mean $S\times d_k/2$
and variance of 1. The sensory evidence $x$ is compared to a decision criterion $c$ 
to generate a discrimination response $R$, which is 1, if $x$ exceeds $c$ and -1 else.
To generate confidence, it is assumed that the confidence variable $y$ is compared to another
set of criteria $\theta_{R,i}, i=1,2,...,L-1$, depending on the
discrimination response $R$ to produce a $L$-step discrete confidence response. The different models
vary in how $y$ is generated (see below). 
The following parameters are shared between all models:

- sensitivity parameters $d_1, ..., d_K$ ($K$: number of difficulty levels),
- decision criterion $c$, 
- confidence criterion $\theta_{-1,1}, ..., \theta_{-1,L-1}, 
\theta_{1,1},  ,...,\theta_{1,L-1}$ ($L$: number of confidence categories available for confidence ratings).

### Signal detection rating model (SDT)
According to SDT, the same sample of sensory evidence is used to generate response and confidence, i.e.,
$y=x$. The confidence criteria associated with $R=-1$ are more negative than the 
decision criterion $c$, whereas the confidence criteria associated with $R=1$ are more positive than $c$.

### Gaussian noise model (GN)
Conceptually, the Gaussian noise model reflects the idea that confidence is informed by the same sensory evidence
as the task decision, but confidence is affected by additive Gaussian noise.
According to GN, $y$ is subject to additive noise and assumed to be normally distributed
around the decision evidence value $x$ with a standard deviation $\sigma$, 
which is an additional free parameter.

### Weighted evidence and visibility model (WEV)
Conceptually, the WEV model reflects the idea that the observer combines evidence about decision-relevant features 
of the stimulus with the strength of evidence about choice-irrelevant features 
to generate confidence. For this purpose, WEV assumes that $y$ is normally 
distributed with a mean of $(1-w)\times x+w \times d_k\times R$ and standard deviation $\sigma$.
The standard deviation quantifies the amount of unsystematic variability
contributing to confidence judgments but not to the discrimination judgments.
The parameter $w$ represents the weight that is put on the choice-irrelevant
features in the confidence judgment. The parameters $w$ and $\sigma$ are free parameters in
addition to the set of shared parameters.

### Post-decisional accumulation model (PDA)
PDA represents the idea of on-going information accumulation after the
discrimination choice. The parameter $a$ indicates the amount of additional
accumulation. The confidence variable is normally distributed with mean
$x+S\times d_k\times a$ and variance $a$. The parameter $a$ is fitted in addition to the shared
parameters.

### Independent Gaussian model (IG)
According to IG, the information used for confidence judgments is generated independently from the sensory evidence used for the task decision. 
For this purpose, it is assumed that $y$ is sampled independently from $x$. 
The variable $y$ is normally distributed with a mean of $a\times d_k$ and variance
of 1. The additional parameter $m$ represents the amount of information available for confidence judgment
relative to amount of evidence available for the discrimination decision and can 
be smaller as well as greater than 1.

### Independent truncated Gaussian model: HMetad-Version (ITGc)
Conceptually, the two ITG models just as IG are based on the idea that the information used for confidence judgments is generated independently from the sensory evidence used for the task decision. However, in contrast to IG,  the two ITG models also reflect a form of confirmation bias in so far as 
it is not possible to collect information that contradicts the original decision. 
According to the version of ITG consistent with the HMetad-method 
(Fleming, 2017), $y$ is sampled independently
from $x$ from a truncated Gaussian distribution with a location parameter
of $S\times d_k \times m/2$ and a scale parameter of 1. The Gaussian distribution of $y$
is truncated in a way that it is impossible to sample evidence that contradicts
the original decision: If $R = -1$, the distribution is truncated to the
right of $c$. If $R = 1$, the distribution is truncated to the left
of $c$. The additional parameter $m$ represents metacognitive efficiency,
i.e., the amount of information available for confidence judgments relative to
amount of evidence available for discrimination decisions and  can be smaller
as well as greater than 1.

### Independent truncated Gaussian model: Meta-d'-Version (ITGcm)
According to the version of the ITG consistent with the original meta-d' 
method (Maniscalco & Lau, 2012, 2014),
$y$ is sampled independently from $x$ from a truncated Gaussian distribution with a location parameter
of $S\times d_k \times m/2$ and a scale parameter of 1. If $R = -1$, 
the distribution is truncated to the right of $m\times c$. If $R = 1$, 
the distribution is truncated to the left of  $m\times c$. The additional parameter $m$ 
represents metacognitive efficiency, i.e., the amount of information available for confidence judgments relative to
amount of evidence available for the discrimination decision and  can be smaller
as well as greater than 1.

### Logistic noise model (logN)
According to logN, the same sample of sensory evidence is used to generate response and confidence, i.e.,
$y=x$ just as in SDT. However, according to logN, the confidence criteria
are not assumed to be constant, but instead they are affected by noise drawn from
a lognormal distribution. In each trial, $\theta_{-1,i}$ is given
by $c -  \epsilon_i$. Likewise, $\theta_{1,i}$ is given by
$c + \epsilon_i$. The noise $\epsilon_i$ is drawn from a lognormal distribution with
the location parameter $\mu_{R,i} = \log(\left| \mu_{\theta_{R,i}} - c\right|)- 0.5 \times \sigma^{2}$, 
and scale parameter $\sigma$. $\sigma$ is a free parameter designed to
quantify metacognitive ability. It is assumed that the criterion noise is perfectly
correlated across confidence criteria, ensuring that the confidence criteria
are always perfectly ordered. Because $\theta_{-1,1}$, ..., $\theta_{-1,L-1}$,
$\theta_{1,1}$, ..., $\theta_{1,L-1}$ change from trial to trial, they are not estimated
as free parameters. Instead, we estimate the means of the confidence criteria, i.e., $\mu_{\theta_{-1,1}}, ...,
\mu_{\theta_{-1,L-1}}, \mu_{\theta_{1,1}}, ...  \mu_{\theta_{1,L-1}}$,
as free parameters.

### Logistic weighted evidence and visibility model (logWEV)
The logWEV model is a combination of logN and WEV proposed by .
Conceptually, logWEV assumes that the observer combines evidence about decision-relevant features of the stimulus with the strength of evidence about choice-irrelevant features. The model also assumes that noise affecting the confidence decision variable is lognormal.
According to logWEV, the confidence decision variable is $y$ is equal to
R &times; y'. The variable y' is sampled from a lognormal distribution with a location parameter of $(1-w)\times x\times R + w \times d_k$ and a scale parameter of $\sigma$. The parameter $\sigma$ quantifies the amount of unsystematic variability
contributing to confidence judgments but not to the discrimination judgments.
The parameter $w$ represents the weight that is put on the choice-irrelevant
features in the confidence judgment. The parameters $w$ and $\sigma$ are free parameters. 

## Measures of metacognition

### Meta-d'/d'
The conceptual idea of meta-d′ is to quantify metacognition in terms of sensitivity
in a hypothetical signal detection rating model describing the primary task, 
under the assumption that participants had perfect access to the sensory evidence
and were perfectly consistent in placing their confidence criteria (Maniscalco & Lau, 2012, 2014). Using a signal detection model describing the primary task to quantify metacognition, it allows a direct comparison between metacognitive accuracy and discrimination performancembecause both are measured on the same scale. Meta-d′ can be compared against the estimate of the distance between the two stimulus distributions estimated from discrimination responses, which is referred to as d′: If meta-d′ equals d′, it means that metacognitive accuracy is exactly as good as expected from discrimination performance. If meta-d′ is lower than d′, it means that metacognitive accuracy is not optimal. It can be shown that the implicit model of confidence underlying the meta-d'/d' method is identical to different versions of the independent truncated Gaussian model (Rausch et al., 2023), depending on whether the original model specification by Maniscalco and Lau (2012) or alternatively the specification by Fleming (2017) is used. We strongly recommend to test whether the independent truncated Gaussian models are adequate descriptions of the data before quantifying metacognitive efficiency  with meta-d′/d′.

### Information-theoretic measures of metacognition

Dayan (2023) proposed several measures of metacognition based on quantities of information theory. 

- Meta-I is a measure of metacognitive sensitivity defined as the mutual information between the confidence and accuracy and is calculated as the transmitted information minus the minimal information given the accuracy:

$$meta-I = I(Y; \hat{Y}, C) - I(Y; \hat{Y})$$ 
This is equivalent to Dayan's formulation where meta-I is the information that confidences transmit about the correctness of a response: 

$$meta-I = I(Y = \hat{Y}; C)$$
- Meta-$I_{1}^{r}$ is meta-I  normalized by the value of meta-I expected assuming 
a signal detection model (Green & Swets, 1966) with Gaussian noise, based on calculating the sensitivity index d':

$$meta-I_{1}^{r} = meta-I / meta-I(d')$$
- Meta-$I_{2}^{r}$ is meta-I normalized by its theoretical upper bound, which is the information entropy of accuracy,  $H(Y = \hat{Y})$:

$$meta-I_{2}^{r} = meta-I / H(Y = \hat{Y})$$

Notably, Dayan (2023) pointed out that a liberal or conservative use of the confidence levels will affected the mutual information and thus all information-theoretic measures of metacognition.

In addition to Dayan's measures, Meyen et al. (submitted) suggested an
additional measure that normalizes the Meta-I by the range of possible values it can take. This required deriving lower and upper bounds of the transmitted information given a participant's accuracy.

$$RMI = \frac{meta-I}{\max_{\text{accuracy}}\{meta-I\}}$$

As these measures are prone to estimation bias, the package offers a simple
bias reduction mechanism in which the observed frequencies of
stimulus-response combinations are taken as the underlying probability
distribution. From this, Monte-Carlo simulations are conducted to estimate
and subtract the bias from these measures. Note that there is probably no way to remove this bias completely.

## Installation

The latest released version of the package is available on CRAN via

```         
install.packages("statConfR")
```

The easiest way to install the development version is using `devtools`
and install from GitHub:

```         
devtools::install_github("ManuelRausch/StatConfR")
```
<!-- without any dots, the code chunk will be shown, but not executed -->

## Usage

### Example data set

The package includes a demo data set from a masked orientation
discrimination task with confidence judgments (Hellmann et al., 2023,
Exp. 1).

```{r}
library(statConfR)
data("MaskOri")
head(MaskOri)
```

### Fitting

The function `fitConfModels` allows the user to fit several confidence
models separately to the data of each participant. 
The data should be provided via the argument `.data` in the form of a data.frame 
object with the following variables in separate columns:

-   stimulus (factor with 2 levels): The property of the stimulus which
    defines which response is correct
-   diffCond (factor): The experimental manipulation that is expected to
    affect discrimination sensitivity
-   correct (0-1): Indicating whether the choice was correct (1) or
    incorrect(0).
-   rating (factor): A discrete variable encoding the decision
    confidence (high: very confident; low: less confident)
-   participant (integer): giving the subject ID.
The argument `model` is used to specify which model should be fitted, with 'WEV', 'SDT', 'GN', 'PDA', 'IG', 'ITGc', 'ITGcm', 'logN', and 'logWEV' as available options. If model="all" (default), all implemented models will be fit, although this may take a while. 

Setting the optional argument `.parallel=TRUE` parallizes model fitting over all but 1 available core. Note that the fitting procedure takes may take a considerable amount of time, 
especially when there are multiple models, several difficulty conditions, and/or  several confidence categories.  For example, if there are five difficulty conditions and five confidence levels, fitting the WEV model to one single participant may take 20-30 minutes on a 2.8GHz CPU. We recommend parallelization to keep the required time tolerable. 

```{r}     
fitted_pars <- fitConfModels(MaskOri, models=c("ITGcm", "WEV"), .parallel = TRUE) 
```

The output is then a data frame with one row for each combination of 
participant and model and separate columns for each estimated parameter as well as for different measures for goodness-of-fit (negative log-likelihood, BIC, AIC and AICy<sub>c</sub>).
These may be used for statistical  model comparisons.

```{r}
head(fitted_pars)
```

It can be seen that the independent truncated Gaussian model is consistently outperformed by the weighted evidence and visibility model, which is why we would not recommend using meta-d′/d′ for this specific task. 

### Visualization 

After obtaining the model fit, it is strongly recommended to visualise the predictions implied by the best-fitting set of parameters and to compare the predictions with the actual data (Palminteri et al., 2017). The `statConfR' package provides the function `plotConfModelFit', which plots the empirically observed distribution of responses and confidence ratings (on the x-axis) as a function of discriminability (in the rows) and stimulus (in the columns) as bars. Superimposed on the empirical data, the plot also shows the prediction of a selected model as dots. The parameters of the model are passed to `plotConfModelFit' by the argument `fitted_pars'

<!-- Stuff where only the code should be shown and executed, but do not show R yapping  -->

```{r, echo=TRUE, results="hide", message=FALSE, warning=FALSE}
PlotFitWEV <- plotConfModelFit(MaskOri, fitted_pars, model="WEV")
PlotFitITGcm <- plotConfModelFit(MaskOri, fitted_pars, model="ITGcm")
```
<!-- Show both the code and the output Figure!  -->

```{r, echo=TRUE, fig.cap = "Observed distribution of accuracy and responses as a function of discriminability and stimulus vs. prediction by the weighted evidence and visibility model"}
PlotFitWEV
```

```{r, echo=TRUE, fig.cap = "Observed distribution of accuracy and responses as a function of discriminability and stimulus vs. prediction by the Independent truncated Gaussian model: HMetad-Version (ITGc)"}
PlotFitITGcm
```

### Measuring metacognition

Assuming that the independent truncated Gaussian model provides a decent account of
the data (notably, this is not the case though in the demo data set), the function `fitMetaDprime` can be used to estimate meta-d′/d′ independently for each subject. The arguments `.data` and `.parallel=TRUE` just in the same way the arguments of `fitConfModels`. The argument `model` offers the user the choice between two model specifications, either "ML" to use the original model specification used by Maniscalco and Lau (2012, 2014) or "F" to use the model specification by Fleming (2017)'s Hmetad method.

```{r}        
MetaDs <- fitMetaDprime(data = MaskOri, model="ML", .parallel = TRUE)
head(MetaDs)
```

Information theoretic measures of metacognition can be obtained by the function `estimateMetaI`. It expects the same kind of data.frame as `fitMetaDprime` 
and returns separate estimates of meta-I, Meta-$I_{1}^{r}$, and Meta-$I_{2}^{r}$ for each subject. 
The preferred way to estimate these measures is with bias reduction, but this may take ~ 6 s for each subject. 

```{r}        
metaIMeasures <- estimateMetaI(data = MaskOri, bias_reduction = TRUE)
head(metaIMeasures)
```

For the impatient and for testing purposes, bias_reduction can be turned off to increase computation speed: 

```        
metaIMeasures <- estimateMetaI(data = MaskOri, bias_reduction = FALSE)
```

### Documentation

After installation, the documentation of each function of of `statConfR` can be accessed by typing *?functionname* into the console. 

## Contributing to the package 
The package is under active development. We are planning to implement new models of decision confidence when they are published. Please feel free to [contact us](malto::manuel.rausch@ku.de) to suggest new models to implement in in the package, or to volunteer adding additional models. 

### Instruction for implementing custom models of decision confidence 
__Only recommended for users with experience in cognitive modelling!__
For readers who want to use our open code to implement models of confidence themselves, the following steps need to be taken: 

- Derive the likelihood of a binary response ($R=-1, 1$) and a specific level of confidence ($C=1,...K$) according to the custom model and a set of parameters, given the binary stimulus ($S=-1, 1$), i.e. $P(R, C | S)$. 
- Use one of the files named 'int_ll*model*.R' from the package sources and adapt the likelihood function according to your model. According to our convention, name the new file a 'int_ll*yourmodelname*.R'. Note that all parameters are fitted on the reals, i.e. positive parameters should be transformed outside the log-likelihood function (e.g. using the logarithm) and back-transformed within the log-likelihood function (e.g. using the exponential).
- Use one of the files 'int_fit*model*.R'  from the package sources and adapt the fitting function to reflect the new model. 
  - The initial grid used during the grid search should include a plausible range of all  parameters of your model. 
  - If applicable, the parameters of the initial grid needs be transformed so the parameter vector for optimization is real-valued).
  - The optimization routine should call the new log-likelihood function.
  - If applicable, the parameter vector i obtained during optimization needs to back-transformation for the the output object `res`. 
  - Name the new file according to the convention 'int_fit*yourmodelname*.R'.
- Add your model and fitting-functions to the high-level functions `fitConf` and `fitConfModels`. 
- Add a simulation function in the file 'int_simulateConf.R' which uses the same 
  structure as the other functions but adapt the likelihood of the responses.
  

## Contact

For comments, bug reports, and feature suggestions please feel free to write to either
[manuel.rausch@hochschule-rhein-waal.de](mailto:manuel.rausch@hochschule-rhein-waal.de)
or [sebastian.hellmann\@ku.de](mailto:sebastian.hellmann@ku.de)
or [submit an issue](https://github.com/ManuelRausch/StatConfR/issues).

## References
- Daunizeau, J., Adam, V., & Rigoux, L. (2014). Vba: A probabilistic treatment of nonlinear models for neurobiological and behavioural data. \emph{PLOS Computational Biology}, 10(1), e1003441. https://doi.org/10.1371/journal.pcbi.1003441
- Dayan, P. (2023). Metacognitive Information Theory. Open Mind, 7, 392–411. https://doi.org/10.1162/opmi_a_00091
- Fleming, S. M. (2017). HMeta-d: Hierarchical Bayesian estimation of metacognitive efficiency from confidence ratings. Neuroscience of Consciousness, 1, 1–14. https://doi.org/10.1093/nc/nix007
- Green, D. M., & Swets, J. A. (1966). Signal detection theory and psychophysics. Wiley.
- Hellmann, S., Zehetleitner, M., & Rausch, M. (2023). Simultaneous modeling of choice, confidence, and response time in visual perception. Psychological Review, 130(6), 1521–1543. https://doi.org/10.1037/rev0000411
- Maniscalco, B., & Lau, H. (2012). A signal detection theoretic method for estimating metacognitive sensitivity from confidence ratings. Consciousness and Cognition, 21(1), 422–430. https://doi.org/10.1016/j.concog.2011.09.021
- Maniscalco, B., & Lau, H. (2016). The signal processing architecture underlying subjective reports of sensory awareness. Neuroscience of Consciousness, 1, 1–17. https://doi.org/10.1093/nc/niw002
- Maniscalco, B., & Lau, H. C. (2014). Signal Detection Theory Analysis of Type 1 and Type 2 Data: Meta-d, Response- Specific Meta-d, and the Unequal Variance SDT Model. In S. M. Fleming & C. D. Frith (Eds.), The Cognitive Neuroscience of Metacognition (pp. 25–66). Springer. https://doi.org/10.1007/978-3-642-45190-4_3
- Palminteri, S., Wyart, V., & Koechlin, E. (2017). The importance of falsification in computational cognitive modeling. Trends in Cognitive Sciences, 21(6), 425–433. https://doi.org/10.1016/j.tics.2017.03.011
- Rausch, M., Hellmann, S., & Zehetleitner, M. (2018). Confidence in masked orientation judgments is informed by both evidence and visibility. Attention, Perception, and Psychophysics, 80(1), 134–154. https://doi.org/10.3758/s13414-017-1431-5
- Rausch, M., & Zehetleitner, M. (2017). Should metacognition be measured by logistic regression? Consciousness and Cognition, 49, 291–312. https://doi.org/10.1016/j.concog.2017.02.007
- Rigoux, L., Stephan, K. E., Friston, K. J., & Daunizeau, J. (2014). Bayesian model selection for group studies - revisited. \emph{NeuroImage}, 84, 971–985. https://doi.org/10.1016/j.neuroimage.2013.08.065
- Shekhar, M., & Rahnev, D. (2021). The Nature of Metacognitive Inefficiency in Perceptual Decision Making. Psychological Review, 128(1), 45–70. https://doi.org/10.1037/rev0000249
- Shekhar, M., & Rahnev, D. (2024). How Do Humans Give Conﬁdence? A Comprehensive Comparison of Process Models of Perceptual Metacognition. Journal of Experimental Psychology: General, 153(3), 656–688. https://doi.org/10.1037/xge0001524

