---
title: "README"
author: "Sebastian Hellmann and Manuel Rausch"
date: "2023-09-15"
output:
  md_document:
    variant: gfm
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
gitrep <- "https://github.com/ManuelRausch/StatConfR/tree"
gitbranch <- "main/"
```

# statConfR: Models of Decision Confidence and Metacognition

This package includes functions to fit meta-d´/d´ as well as several
static models of decision making and confidence derived from signal detection theory to test the
assumptions underlying meta-d´/d´ and which may serve as an alternative when the assumptions of meta-d´/d´ do not hold. Currently, the following models are included:

-   Signal Detection Rating Model

-   Gaussian Noise Model

-   Weighted Evidence and Visibility Model

-   Post-decisional Accumulation Model

-   Independent Gaussian model

-   Independent Truncated Gaussian Model (the model underlying the
    meta-d´/d´ method)

See Rausch et al. (2023) for more detail about the models.

## Installation

For the current development version, the easiest way of installation is
using `devtools` and install from GitHub:

```         
devtools::install_github("ManuelRausch/StatConfR")
```

The latest released version of the package will soon be available on
CRAN via

``` r
install.packages("statConfR")
```

## Usage

### Data structure

The package includes a demo data set from a masked orientation
discrimination task with confidence judgments.

```{r}
library(statConfR)
data("MaskOri")
head(MaskOri)
```

Data should be in the form of a data.frame object columns for following
variables:

-   stimulus (factor with 2 levels): The property of the stimulus which
    defines which response is correct
-   condition (factor): The experimental manipulation that is expected
    to affect discrimination sensitivity (only used for fitting
    confidence models)
-   correct (0-1): Indicating whether the choice was correct (1) or
    incorrect(0).
-   rating (factor): A discrete variable encoding the decision
    confidence (high: very confident; low: less confident)
-   sbj (integer): giving the subject ID.

```{r}
MaskOri$condition <- MaskOri$diffCond
head(MaskOri)
```

### Fitting

It is strongly recommended that if metacognitive efficiency is to be
measured using the meta-d'/d' method that researchers fist determine
whether the Independent Truncated Gaussian Model, the confidence model
implied by the meta-d'/d' method, is an adequate description of the
data. Using the function fitConfModel, we can fit several confidence
models to the data of each participant. The argument
`.parallel=TRUE`allows for parallelization over all but one available
core.

``` r
 fitted_pars <- fitConfModels(MaskOri, models=c("SDT", "WEV"), .parallel = TRUE) 
```

This parallelizes the fitting process over participant-model
combinations. The output is then a data frame with one row for each
participant-model combination and columns for parameters and measures
for model performance (negative log-likelihood, BIC, AIC and AICc).
These may be used for quantitative model comparison.

```{r, echo=FALSE}
fitted_pars <- structure(list(model = c("SDT", "WEV", "SDT", "WEV", "SDT", "WEV", 
"SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", 
"WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", 
"SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", 
"WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", 
"SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", 
"WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", 
"SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", "WEV", "SDT", 
"WEV", "SDT", "WEV"), participant = c(1, 1, 2, 2, 3, 3, 4, 4, 
5, 5, 6, 6, 7, 7, 8, 8, 9, 9, 10, 10, 11, 11, 12, 12, 13, 13, 
14, 14, 15, 15, 16, 16, 17, 17, 18, 18, 19, 19, 20, 20, 21, 21, 
22, 22, 23, 23, 24, 24, 25, 25, 26, 26, 27, 27, 28, 28, 29, 29, 
30, 30, 31, 31, 32, 32, 33, 33, 34, 34, 35, 35, 36, 36), negLogLik = c(989.806844654359, 
913.081913018423, 1076.5072058806, 1053.69810875527, 890.299976524937, 
767.722494118762, 1542.16768994555, 1347.2354843231, 1362.89158239783, 
1287.38665943591, 917.321875454359, 840.895557645978, 1151.88251752874, 
1039.64350633853, 1248.17509704189, 1050.09049296071, 1162.00570515158, 
1045.6032329578, 1159.85277656942, 939.822495755436, 1152.37215662276, 
1146.52380941997, 1102.63823377065, 994.134901753055, 1164.17397549339, 
1155.47194461487, 1068.52786624312, 1063.69896452573, 1311.97921994465, 
1181.92858927924, 1463.99416938906, 1241.19948621088, 1587.75813049321, 
1435.94489241206, 762.888678358319, 744.947678173956, 898.071847266369, 
781.548952139715, 1069.45979928833, 1062.04911800501, 1421.84932957052, 
1365.59431520477, 1642.62751419503, 1616.07791873891, 1123.33375019946, 
1082.15614299351, 1239.57514484883, 1104.72050948434, 815.137236069942, 
785.018506205193, 992.734071112081, 925.678628373154, 1361.19296495905, 
1219.91773179505, 1215.35643353453, 1192.29486591689, 1229.13830575971, 
1145.6782231611, 1177.90522684965, 1156.05891796667, 1242.97705880543, 
1148.10371730271, 1225.58170194889, 1090.03322503802, 922.067673221904, 
803.666563648033, 1202.18614343591, 1053.05792910063, 982.619518664301, 
964.973937924523, 1260.37939489443, 1073.14022156417), N = c(960L, 
960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 
960L, 960L, 830L, 830L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 
960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 
960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 
960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 
960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 960L, 
960L, 960L, 960L, 960L, 960L), k = c(11L, 13L, 11L, 13L, 11L, 
13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 
11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 
13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 
11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 
13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 11L, 13L, 
11L, 13L), BIC = c(2055.1499554378, 1915.43395873485, 2228.55067789028, 
2196.66635020853, 1856.13621917895, 1624.71512093553, 3159.87164602018, 
2783.7411013442, 2801.31943092475, 2664.04345156981, 1910.1800170378, 
1771.06124798996, 2379.30130118657, 2168.55714537506, 2570.28587679247, 
2187.5595200317, 2399.54767643224, 2180.4765986136, 2395.24181926792, 
1968.91512420888, 2380.2805793746, 2382.31775153794, 2280.81273367037, 
2077.53993620411, 2403.88421711586, 2400.21402192775, 2212.59199861531, 
2216.66806174947, 2699.49470601838, 2453.12731125648, 3003.52460490719, 
2571.66910511975, 3251.05252711549, 2961.15991752213, 1601.31362284572, 
1579.16548904592, 1871.67996066182, 1652.36803697743, 2214.45586470575, 
2213.36836870803, 2919.23492527012, 2820.45876310754, 3360.79129451913, 
3321.42597017582, 2322.20376652801, 2253.58241868501, 2554.68655582674, 
2298.71115166668, 1705.81073826896, 1659.30714510839, 2061.00440835324, 
1940.62738944431, 2797.92219604718, 2529.10559628811, 2506.24913319815, 
2473.85986453179, 2533.81287764851, 2380.62657902021, 2431.34671982838, 
2401.38796863134, 2561.49038373993, 2385.47756730342, 2526.69967002686, 
2269.33658277403, 1919.67161257289, 1696.60325999407, 2479.90855300091, 
2195.38599089927, 2040.77530345768, 2019.21800854705, 2596.29505591795, 
2235.55057582634), AICc = c(2001.84575681927, 1852.493635762, 
2175.24647927175, 2133.72602723569, 1802.83202056042, 1561.77479796268, 
3106.56744740165, 2720.80077837135, 2748.01523230622, 2601.10312859697, 
1856.87581841927, 1708.12092501711, 2325.99710256803, 2105.61682240222, 
2518.61914273903, 2126.5633388626, 2346.24347781371, 2117.53627564076, 
2341.93762064939, 1905.97480123603, 2326.97638075606, 2319.37742856509, 
2227.50853505184, 2014.59961323127, 2350.58001849732, 2337.2736989549, 
2159.28779999678, 2153.72773877663, 2646.19050739985, 2390.18698828364, 
2950.22040628866, 2508.72878214691, 3197.74832849696, 2898.21959454929, 
1548.00942422719, 1516.22516607307, 1818.37576204329, 1589.42771400459, 
2161.15166608722, 2150.42804573518, 2865.93072665158, 2757.51844013469, 
3307.4870959006, 3258.48564720298, 2268.89956790948, 2190.64209571217, 
2501.38235720821, 2235.77082869384, 1652.50653965043, 1596.36682213554, 
2007.70020973471, 1877.68706647147, 2744.61799742864, 2466.16527331526, 
2452.94493457961, 2410.91954155894, 2480.50867902997, 2317.68625604737, 
2378.04252120984, 2338.4476456585, 2508.1861851214, 2322.53724433058, 
2473.39547140833, 2206.39625980119, 1866.36741395436, 1633.66293702122, 
2426.60435438237, 2132.44566792642, 1987.47110483915, 1956.2776855742, 
2542.99085729942, 2172.61025285349), AIC = c(2001.61368930872, 
1852.16382603685, 2175.0144117612, 2133.39621751053, 1802.59995304987, 
1561.44498823752, 3106.3353798911, 2720.4709686462, 2747.78316479567, 
2600.77331887181, 1856.64375090872, 1707.79111529196, 2325.76503505748, 
2105.28701267706, 2518.35019408378, 2126.18098592142, 2346.01141030316, 
2117.2064659156, 2341.70555313884, 1905.64499151087, 2326.74431324551, 
2319.04761883993, 2227.27646754129, 2014.26980350611, 2350.34795098678, 
2336.94388922974, 2159.05573248623, 2153.39792905147, 2645.9584398893, 
2389.85717855848, 2949.98833877811, 2508.39897242175, 3197.51626098641, 
2897.88978482413, 1547.77735671664, 1515.89535634791, 1818.14369453274, 
1589.09790427943, 2160.91959857667, 2150.09823601002, 2865.69865914103, 
2757.18863040953, 3307.25502839005, 3258.15583747782, 2268.66750039893, 
2190.31228598701, 2501.15028969766, 2235.44101896868, 1652.27447213988, 
1596.03701241039, 2007.46814222416, 1877.35725674631, 2744.3859299181, 
2465.83546359011, 2452.71286706907, 2410.58973183379, 2480.27661151943, 
2317.35644632221, 2377.8104536993, 2338.11783593334, 2507.95411761085, 
2322.20743460542, 2473.16340389778, 2206.06645007603, 1866.13534644381, 
1633.33312729607, 2426.37228687182, 2132.11585820127, 1987.2390373286, 
1955.94787584905, 2542.75878978887, 2172.28044312833), d1 = c(0.150713840702265, 
0.138541992497106, 0.242773155633049, 0.37470202088794, 0.0160943292873731, 
0.00244278043759237, 0.1136495828561, -0.0880232454706361, -0.0867599729712157, 
-0.103675675475424, 0.212925807363546, 0.399023261956908, 0.0783712927183564, 
-0.0696630553208628, -0.0328720479063638, -0.20328282068976, 
0.0604792741415441, 0.0588201759313646, 0.065145418869179, 0.0463672046584133, 
0.54884979167218, 0.597137313482079, 0.207828707870216, 0.594779460392217, 
0.18005416216454, 0.232807667274392, 0.727746151287307, 0.883618961595971, 
-0.0779427425278179, -0.237288391194024, 0.0323432861616166, 
0.0915132762137052, 0.0395359965785297, -0.0858362810213649, 
-0.0660299734211784, -0.0924377705537742, 0.308439102897019, 
0.512995998214726, 0.0240540896743552, 0.0903009278204649, -0.0136654499983109, 
-0.0938776620687124, 0.0380860256814059, 0.0053865569327196, 
0.145197874696882, 0.273358243778637, 0.276724407440858, 0.337936258129005, 
0.542116585773805, 0.767483310326329, 0.183029463982222, 0.22288908890138, 
-0.0886412062483802, -0.502065365184728, 0.53451932212813, 0.87086111378948, 
0.255281842354408, 0.371663688389825, 0.257048914927863, 0.530453971545555, 
0.179569376810233, 0.152129816499047, 0.0604708637409885, -0.0897555938563303, 
0.0108251218876531, -0.032525285382168, -0.0497925232707996, 
-0.318023560769132, 0.314474683979523, 0.193791535043305, -0.0522409926509587, 
-0.167242276221158), d2 = c(1.16229923539805, 1.44991336503167, 
1.41386446964372, 1.44420482248688, 0.140493809705336, 0.407598221011013, 
0.310332845134505, 0.74082057501481, 0.213420317688283, 0.229488775695156, 
1.28765536225901, 1.49458779868544, -0.177823629824188, -0.123850205129803, 
0.121671701434239, 0.257717235231953, 0.291516236121905, 0.497082162907943, 
0.292128395217934, 0.533975626938575, 1.929303997313, 1.95906784279179, 
1.33761040979721, 1.7132334944555, 0.689003260243324, 0.992861422579516, 
2.62137029607618, 2.82071954469085, -0.0174049876085645, -0.0570834587341738, 
0.347809728810492, 0.447791125103373, 0.0349921888617545, 0.125601877271279, 
2.80206537124732, 2.58054661089962, 0.536751922790548, 0.784590363578053, 
0.741490108961578, 0.749198448928945, 0.323418389017221, 0.603728754623058, 
0.318016272295883, 0.307009573220564, 1.02663080257776, 1.11553598039618, 
0.57840994031023, 0.785696437996745, 3.11954340068947, 3.03914016912602, 
0.320287267595874, 0.376647512218371, -0.169282403504683, -0.147107469324142, 
2.25465657712097, 2.37105964494075, 0.367932735836438, 0.584020251944118, 
1.18288092344232, 1.41242028278719, 0.640392268438897, 1.01042120699047, 
0.176248415764694, 0.296772039394003, 0.420685358121356, 0.627147706362697, 
0.107001110474169, 0.330572848059471, 0.677339497661569, 1.02730480971798, 
0.00441428855173975, -0.13012061044839), d3 = c(5.86672779527598, 
4.23324945532841, 4.92586046515988, 4.07391929782287, 2.94772474747739, 
3.20650307475054, 1.74398630600026, 1.76266314662682, 2.14390612806672, 
2.37305543549854, 4.40642161397509, 3.40678186474995, 1.25007939518623, 
1.5905469299652, 2.05242493289332, 1.91662441917313, 1.92275028643378, 
2.0232076670498, 2.24537737419141, 2.31744643853814, 4.41018265993085, 
4.09723739565908, 3.88202483175032, 3.45488073757055, 3.06922167811152, 
3.90462545089351, 5.75479796143903, 5.49910755720531, 1.29723453044741, 
1.54179608600946, 1.17224698847603, 1.62866700604107, 0.842537970088249, 
1.22007713272453, 7.03173896729844, 5.64060190881236, 3.90736662885707, 
2.87554890446631, 4.24249845684069, 3.71153970907723, 1.78156950848646, 
1.9664427553227, 1.17743245127692, 1.44746326097919, 4.26438472887423, 
3.39926374476107, 1.33274971480231, 1.67518563618671, 5.88992909584795, 
4.67925372194701, 3.40414699363089, 2.60072987350241, 1.07889121046973, 
1.41776795176489, 4.61039065379267, 4.11357660175089, 2.11527671108533, 
2.06207610624963, 2.20420955938701, 2.40118912445419, 2.9338346686024, 
2.52874136227192, 1.60099622913109, 1.58246905062863, 3.96408445451527, 
3.02625439371347, 2.30652175424617, 2.10543117248538, 6.02343155296529, 
4.38101757311111, 1.45700038943829, 1.64850370861961), d4 = c(8.62539593573474, 
5.50672126132543, 7.6918361767904, 5.97199643552606, 7.0869754896812, 
5.92601248188455, 4.63711300915415, 2.72910197527018, 5.16546050851182, 
3.80252112499265, 5.88974347303902, 3.99628123694469, 3.4991065974703, 
2.71403469442601, 4.50637803578958, 2.79697542046017, 4.30393602664759, 
3.12918244952418, 6.71396300370228, 3.82819081454796, 7.57476053212253, 
6.76642974034756, 8.79939148594023, 6.25600829620929, 4.73468881165036, 
5.88274053090394, 7.1229030635091, 6.67497609859386, 3.46095507449844, 
2.16334806215014, 4.64487646424577, 2.53828921719507, 4.25763736704446, 
2.87201843669439, 6.84950207079694, 5.57724889944041, 10.2096534713426, 
5.73528982475937, 8.33433695753581, 7.01859549184896, 3.66758796663653, 
3.005279354137, 3.61696815515319, 3.04203456456626, 7.6079181036788, 
5.79582116661863, 6.56172235456445, 3.15085407486311, 6.6846813807782, 
5.29871286650265, 8.01612572698049, 4.78508316873425, 5.30173157470097, 
2.91975888271235, 6.26684528264362, 5.45690767337529, 6.14942607847974, 
3.87062589389098, 5.06346122440535, 4.65406620303942, 7.22555107850457, 
4.41924365512823, 6.70714708678752, 3.77427539198824, 6.47620623222963, 
4.02682511463261, 5.63387438593021, 3.20707168255937, 8.23447129757913, 
5.80157273447577, 6.41903872274964, 3.15889442210196), theta = c(0.385755909452848, 
0.398778781964313, 0.0278424767295187, 0.0259552453161252, -1.26647185606533, 
-1.33589281792371, -0.0983293204260389, -0.090273549037261, -0.764350087757061, 
-0.784249681612538, -0.125539986255414, -0.0779046340371391, 
0.480301346868824, 0.472817482612408, 0.00612461144714488, 0.00539281211818904, 
-0.551202429682567, -0.560095981844706, -0.658998689663107, -0.627068690400079, 
-0.0257929948883865, -0.0151828824418331, -0.655940145660236, 
-0.678561032282079, -0.886641725651531, -1.03515451594596, 0.391188343423407, 
0.441712446173794, -0.485594316111663, -0.448294186773515, 0.197946997080029, 
0.190930089274182, -0.466681683297304, -0.46631204458053, -0.425294780853032, 
-0.391782002977956, -0.320449359703094, -0.306243496213014, -0.565899360854205, 
-0.558334588655527, -0.703308518042977, -0.697199359756174, 0.083769373107608, 
0.0884361442692118, -0.0148941264253255, -0.0122695438459491, 
0.127010492103701, 0.114243793203813, 0.343388405611671, 0.348136861647931, 
0.123069987265104, 0.144067152290376, -0.146107259391941, -0.131458683919094, 
-0.560411729391698, -0.563865346052207, 0.0188966359499249, -0.003920801746103, 
-0.420554589715601, -0.441404921525925, -0.454213599782394, -0.428806073860464, 
0.112116987741389, 0.11677606140525, -0.210975586572536, -0.17128414004898, 
-0.106063788797856, -0.0844132045620606, -0.26925233023303, -0.250502349982006, 
0.122423389077554, 0.111709280382454), w = c(NA, 0.414761885471073, 
NA, 0.233263711135156, NA, 0.436068779301523, NA, 0.542682764050194, 
NA, 0.385917970129683, NA, 0.426702081297893, NA, 0.671585177202471, 
NA, 0.535709738802725, NA, 0.510217389314206, NA, 0.675401379863473, 
NA, 0.134767261998615, NA, 0.607483392688422, NA, 0.000522085545199569, 
NA, 0.251416867378461, NA, 0.446567197054045, NA, 0.63517915665458, 
NA, 0.665540400953292, NA, 0.182854162286665, NA, 0.57094175911117, 
NA, 0.132045991847071, NA, 0.389645144407971, NA, 0.322033216549472, 
NA, 0.333824082051445, NA, 0.605587394859002, NA, 0.352762373166057, 
NA, 0.348921468298657, NA, 0.53386176009837, NA, 0.29792244315922, 
NA, 0.571522587604453, NA, 0.390107493488682, NA, 0.528793740288083, 
NA, 0.558573500964107, NA, 0.404461797563143, NA, 0.487782514392174, 
NA, 0.356044148328719, NA, 0.566993943298006), sigma = c(NA, 
0.512399891390972, NA, 0.408634132562893, NA, 0.687549970821055, 
NA, 0.370979209508551, NA, 0.599576951678603, NA, 0.426981981206835, 
NA, 0.857916206995298, NA, 0.380424535629457, NA, 0.553715787695801, 
NA, 0.482746674157497, NA, 0.439324812498967, NA, 0.738875192206419, 
NA, 1.10673651376955, NA, 0.808485136616266, NA, 0.239714589453548, 
NA, 0.376080764940219, NA, 0.6463982162105, NA, 0.328015990723485, 
NA, 0.492837074574317, NA, 0.295953189187999, NA, 0.597272106870916, 
NA, 0.715801223303761, NA, 0.539190864018748, NA, 0.331963551399836, 
NA, 0.608103046864028, NA, 0.0375867691130402, NA, 0.484955026718985, 
NA, 0.635954821013192, NA, 0.612473987832079, NA, 0.951087021346356, 
NA, 0.556115508702939, NA, 0.466520266670703, NA, 0.336303624784757, 
NA, 0.421345745527519, NA, 0.597386787271095, NA, 0.395496319035518
), cA1 = c(-1.76340595477229, -0.82076724660998, -2.10614543018389, 
-1.47452434300732, -2.28680674030665, -1.90567502407084, -2.03454371501719, 
-1.22629100302106, -1.49458558943929, -0.981848733565906, -1.26818870033911, 
-0.676639739363507, -5.44335610001409, -4.97667502411756, -2.29305965358962, 
-1.47956020128963, -1.23198718745121, -0.67629231772676, -2.79582557734215, 
-1.77621858436417, -1.83209449126261, -1.49843368966342, -2.46912890248288, 
-1.36319810925356, -1.35370483283875, -1.32410304008043, -2.60852980973843, 
-2.15847169086053, -1.57913338850833, -1.18830814081394, -2.43912566432139, 
-1.40784478462721, -1.31332195964217, -0.588605512116729, -1.64029731113136, 
-1.07284676693135, -3.20748419641158, -1.74900804026308, -2.62835008954678, 
-2.16774198227495, -1.89380200274677, -1.60512051522124, -0.964831220116622, 
-0.760904717908809, -1.67220070737751, -0.919857644770681, -2.50755328694709, 
-1.25518489101325, -1.38627850578312, -0.584346454167642, -3.76841614225674, 
-2.36701954629146, -1.96048652771282, -1.33530810855355, -2.4834249801648, 
-1.93693258533406, -2.12757218599027, -1.32278613873214, -1.40666049357006, 
-0.998238238480247, -2.15309706264256, -1.28096299089893, -2.07373523439259, 
-1.29602507323307, -1.31391608852679, -0.798741826242555, -1.38171683570694, 
-0.860829593904868, -1.94638611919467, -0.999711909063603, -2.32958610131547, 
-1.33991216855537), cA2 = c(-1.11696101974791, -0.242322467402805, 
-1.54857032269223, -0.967136837277126, -1.78780670448711, -1.23416073955202, 
-1.1559581931487, -0.585779491924463, -1.2285521718857, -0.552701006601234, 
-1.26811560657394, -0.676517260544409, -2.56185871453264, -2.32104348717708, 
-1.40257269437971, -0.817000187782491, -1.19763781328201, -0.625340220424436, 
-1.70799199480818, -1.01228824792056, -1.05475753912422, -0.743013900335056, 
-1.48583365824118, -0.218114007714942, -1.21497886579478, -0.98457333201299, 
-1.51472911340725, -0.964478429763883, -1.26974573634855, -0.990410193467559, 
-1.47193779379892, -0.819374993039169, -0.981902504267361, -0.0707963074129448, 
-1.49100708967628, -0.940866670428942, -2.52497004340616, -1.19362800102613, 
-1.98423428405552, -1.57261021491044, -1.31788448654669, -1.00223233300007, 
-0.686374959691877, -0.422543673759392, -1.18511151878222, -0.444783907381211, 
-1.72573136865703, -0.765450977200054, -0.917933396942235, -0.10255758123661, 
-2.96740934353592, -1.83141117619984, -1.40363599525766, -0.882960240619568, 
-1.77152019324168, -1.21343093676252, -1.16504527705015, -0.373447296040246, 
-0.642970387291233, 0.438773042762899, -1.31954206861941, -0.52615027301195, 
-0.942234631815685, -0.28738584133729, -1.11989623282892, -0.595874221373458, 
-1.27202324206742, -0.756167105190736, -1.54410599069954, -0.613713036027049, 
-1.63525506268986, -0.884247978112285), cA3 = c(-0.125426048580949, 
0.828972953073924, -0.909721117712021, -0.385933719058654, -1.35670379046407, 
-0.0761942076809054, -0.375855484492408, 0.274560667802693, -0.905770044629741, 
0.112854464195826, -0.776830026093783, -0.229794447228347, -0.832302126051614, 
-0.400386027001757, -0.976351657519291, -0.393810742276831, -1.05813271811894, 
-0.421875491724098, -1.10665471721819, -0.106560252589737, -0.32854332451099, 
0.0672655557491193, -0.795425513324416, 1.16962903620189, -0.912970992772067, 
0.646482079380039, 0.0300225755665317, 0.977451790631538, -0.884625410649562, 
-0.651256037346372, -0.685909260502863, -0.127161701465275, -0.655057924771073, 
0.58222804395477, -1.07088294104967, -0.562129146002307, -1.47698467262544, 
-0.262003422460817, -1.32917197960084, -0.973448643557498, -0.994346348227608, 
-0.520320219240658, -0.326230567156557, 0.109233584827752, -0.261774004529464, 
0.664393153563925, -1.04967052609546, -0.289335127793152, 0.268018574424538, 
1.58390190451715, -1.86165052630309, -1.05487075146751, -0.846271915211358, 
-0.281724391612968, -1.09577542691366, -0.44763430005973, -0.0265237642830382, 
1.29296135013166, -0.440178648030241, 1.65862294902674, -0.686332032794563, 
0.399869053044363, 0.0520223239309093, 1.15619490464718, -0.91424659534335, 
-0.34693015879225, -0.910361987800253, -0.344141107066154, -0.847848141283355, 
0.151096484768166, -1.03146348876353, -0.355662193784363), cB1 = c(1.02216399395518, 
-0.127256895502679, 0.794431919144025, 0.240448426587239, 0.887542691151107, 
0.28612745913299, 0.441352582658827, -0.157198127879191, -0.392750639265374, 
-0.905476315346184, 1.39655634117865, 0.577725155581914, 1.10306081590396, 
0.363174336501528, 0.852888478500434, 0.301980403837695, 0.643684479685035, 
0.159008127591618, 0.913326952340771, -0.00170754546557261, 0.514839517084559, 
0.153457397387485, 0.385817319083826, -0.868966008814973, -0.634418672021161, 
-1.50775714109563, 0.80491402410481, -0.136015712993464, 1.19621110706906, 
0.665576173109446, 1.04877261416804, 0.295274819125068, -0.162271513280281, 
-0.950724463084392, 0.98236647592937, 0.409131346881463, 1.37010511878912, 
0.129850250631159, 0.828667254205236, 0.466616635596124, 0.722866844001037, 
0.410766353823429, 0.446237544666869, -0.0867573148316765, 0.244038598579574, 
-0.656937587392823, 1.06829251754451, 0.284825549771739, 0.449073878851491, 
-0.964132353737215, 1.52471954331006, 0.775877227252664, 1.04966568855226, 
0.424220109575747, 0.661799332848643, 0.088711477454967, 0.0812478575391518, 
-1.28902317376734, -0.382096367454829, -2.18750473816806, 0.544437905398103, 
-0.290190066360971, 0.254903103447647, -0.879245960883231, 1.48853177249348, 
0.628964713849607, 1.1058710972474, 0.449499551502103, 0.53219620061464, 
-0.356921252122284, 1.01887371035088, 0.288397361394383), cB2 = c(1.76917788549, 
0.6705369813805, 1.54730728495926, 0.927025131112821, 1.71883474866131, 
1.27932750318467, 1.18693833708991, 0.636553843535339, 0.362039978671272, 
-0.0626810602738281, 1.76171120803251, 0.914624800861551, 3.04859746187061, 
2.7271523787507, 1.32846233682382, 0.843964075076859, 0.789760012827119, 
0.306282946944675, 1.57872768839134, 0.793394436483371, 1.35186153768381, 
1.01005392543671, 1.19393051265657, 0.0319370506563798, 0.433718398493076, 
0.49985293058425, 1.89427071705462, 1.32962947964829, 1.60760367755598, 
0.98330606707206, 1.45533104666253, 0.732252796771899, 0.385604625526989, 
-0.249748423542228, 1.39089960821812, 0.795591868237546, 2.71538895311308, 
1.29331306153819, 1.46949015094127, 1.05550976827274, 1.33005841014759, 
1.00074203402603, 0.922505133463409, 0.607677798078035, 1.07088941745917, 
0.34317322425891, 1.86150304813271, 0.869554294786801, 1.17183022090179, 
0.300104204635375, 2.60450866955838, 1.58013439278777, 1.76353593032505, 
1.10413486884669, 1.5131143085592, 0.932728710093448, 0.938965522394267, 
0.166860599512754, 0.0200864566473172, -0.821853519835799, 1.35663449502726, 
0.510925741525695, 1.38132827394226, 0.455351904075246, 1.6353448078487, 
0.773975590127883, 1.5556057421848, 0.854137024281238, 1.31621342648073, 
0.404609806708299, 1.47118669365774, 0.784918211373891), cB3 = c(2.18343884807185, 
1.08643407527867, 1.99551527724245, 1.33912533863266, 2.41081805074028, 
2.01862007685725, 1.78904595955142, 1.07396670334722, 0.949525311133886, 
0.580563443295482, 1.76180030795779, 0.914816500795588, 5.51963087588212, 
5.03798405626049, 2.51920117006928, 1.66843328374522, 0.826668578737165, 
0.344262088482474, 2.76772561751352, 1.68327405362591, 2.08800736488159, 
1.73331910857326, 2.16206882342017, 1.05309983016292, 0.986040347998657, 
1.32369933007972, 2.79378054092596, 2.33677939410002, 1.79140359984688, 
1.11658362729906, 2.18602581080393, 1.25169332440085, 0.943220822644866, 
0.392993413041138, 1.53691744311993, 0.934881115245329, 3.60067768235578, 
2.00127406818836, 2.23987341973388, 1.7899000025734, 1.917754655817, 
1.55793189773017, 1.21059653828026, 0.948809012419589, 1.48302959579682, 
0.753052384702263, 2.5592192476722, 1.28298334260547, 1.59475108519375, 
0.768259373397662, 3.660067600565, 2.28624694539282, 2.36351426884061, 
1.5546302016924, 2.32670623744884, 1.73251650056092, 1.68050808658376, 
0.990777742593266, 1.42499483020447, 1.10475956178564, 2.14388343541165, 
1.22127832924463, 2.46037408824605, 1.4686023449785, 1.76336079955311, 
0.896840764344462, 1.68366832132373, 0.95531324271835, 1.72787273134324, 
0.789215672511124, 2.01062235728567, 1.19780832399231)), class = "data.frame", row.names = c(NA, 
-72L))
```

```{r}
head(fitted_pars)
```

If the Truncated Gaussian model performs best in the model comparison,
metacognitive efficiency can be quantified using meta-d'/d':

``` r
 MetaDs <- fitMetaDprime(MaskOri, model="ML", .parallel = TRUE)
```

## Future development

The package is still under active development. We plan to include more
models from the literature and function to compute the predicted
distribution of responses for given parameter sets. For any suggestions
and contributions, please contact us!

## Contact

For comments, remarks, and questions please contact either
[manuel.rausch\@hochschule-rhein-waal.de](mailto:manuel.rausch@hochschule-rhein-waal.de)
or [sebastian.hellmann\@ku.de](mailto:sebastian.hellmann@ku.de){.email}
or [submit an issue](https://github.com/ManuelRausch/StatConfR/issues).

## References

Rausch, M., Hellmann, S. & Zehetleitner, M. (2023). Measures of
metacognitive efficiency across cognitive models of decision confidence.
[doi: 10.31234/osf.io/kdz34](https://doi.org/10.31234/osf.io/kdz34)
